\newpage
\begin{thebibliography}{99}

    \bibitem{Malik}
    Farhad Malik. \textit{What is Dimension Reduction in Data Science?}. FinTechExplained, \textit{https://www.kdnuggets.com/2019/01/dimension-reduction-data-science.html} (Jan 31, 2019)

    \bibitem{Jaadi}
    Zakaria Jaadi. \textit{A Step-by-Step Explanation of Principal Component Analysis (PCA)}. builtin, \textit{https://builtin.com/data-science/step-step-explanation-principal-component-analysis} (Aug 08, 2022)

    \bibitem{Friedman}
    Hastie T, Tibshirani R, Friedman J. \textit{The Elements of Statistical Learning}.  Data Mining, Inference, and Prediction, second edition, 528-576 (2017)

    \bibitem{Chazal}
    Chazal F, Michel B. \textit{An Introduction to Topological Data Analysis}. Fundamental and Practical Aspects for Data Scientists

    \bibitem{Tharwat}
    Tharwat A. \textit{Independent component analysis: An introduction}. Applied Computing and Informatics, Vol. 17 No. 2, pp. 222-249. \textit{https://doi.org/10.1016/j.aci.2018.08.006} (Aug 06, 2018)

    \bibitem{Aapo}
    Aapo Hyvärinen, Erkki Oja. \textit{Independent Component Analysis: Algorithms and Applications}. Neural Networks, 13(4-5):411-430 (2000) 

    \bibitem{Approx}    
    \textit{R: Interpolation Functions} Psu.edu. approxfun {stats} \textit{https://astrostatistics.psu.edu/su07/R/html/stats/html/approxfun.html} (Mar 22, 2023)
    
    \bibitem{R Stats}   
    \textit{R: The R Stats Package} Psu.edu. \textit{https://astrostatistics.psu.edu/su07/R/html/stats/html/00Index.html} (Mar 22, 2023)

    \bibitem{cocktail}
    Kevin J. P. Woods, Josh H. McDermott. \textit{Schema learning for the cocktail party problem}, pnas.org \textit{https://www.pnas.org/doi/10.1073/pnas.1801614115} (Mar 21 2018)

    \bibitem{fastICA}
    J L Marchini, C Heaton and B D Ripley. \textit{fastICA: FastICA Algorithms to Perform ICA and Projection Pursuit}. R package version 1.2-3. \textit{https://CRAN.R-project.org/package=fastICA} (2021)

    \bibitem{fastICACranR}
    Marchini, J., Heaton, C., Ripley, B., Maintainer, B. and Ripley \textit{Package ‘fastICA’} FastICA Algorithms to Perform ICA and Projection Pursuit \textit{https://cran.r-project.org/web/packages/fastICA/fastICA.pdf } (Mar 22, 2023)

    \bibitem{Varkonyi}
    Tamas Pal, Daniel T. Varkonyi. \textit{Comparison of Dimensionality Reduction Techniques on Audio Signals} Eötvös Loránd University, Department of Data Science and Engineering (2020)

    \bibitem{Pigoli}
    Davide Pigoli, Pantelis Z. Hadjipantelis, John S. Coleman, John A. D. Aston. \textit{The statistical analysis of acoustic phonetic data: exploring differences between spoken Romance languages} Appl. Statist. 67, Part 5, pp. 1103–1145 (2018)

    \bibitem{categorized}
    \textit{What is Dimensionality Reduction? Overview, and Popular Techniques} Simplelearn (Mar 7 2023)

    \bibitem{PCAbiplot}
    \textit{Psychological Statistics.} www.uv.es \textit{ https://www.uv.es/visualstats/vista-frames/help/lecturenotes/lecture13/biplot.html} (Mar 22, 2023)

    \bibitem{R}
    R Core Team. \textit{R: A language and environment for statistical computing}. R Foundation for Statistical Computing, Vienna, Austria. \textit{URL https://www.R-project.org/.} (2021)
    
\end{thebibliography}